Dual host-parasite transcriptomes of apicomplexan Eimeria falciformis and its natural mouse host

### 2016-08-15 manuscript/bmc_article.tex
Up to the enrichment tests

  \section*{A dual transcriptomics experiment}
We performed mRNA sequencing of caecum tissue from mice infected with
the apicomplexan parasite \textit{E. falciformis}. We worked with two
biological replicates for all but two conditions (with one and three
replicates, respecitively). For each replicate sample we enriched
caecum tissue from three infected mice for epithelial cells. We
extracted mRNA and constructed sequencing libraries for different
timepoints after infection and mouse strains which displayed different
immunocompetence in infection trials (see figure x).

In these experiments we analysed RAG-1-/- mice and compared them to
the parental C57BL6 mouse line. RAG knockout mich lack adaptive immune
response du to the absence of mature B and T lymphocytes. Differences
in oocyst output were more pronounced between NMRI and both C57BL6 and
RAG-1-/-. Surprisingly this finding was made both in primary and
challenge infection. Similarly development of intestinal stages as
assayed by quantitative reverse transcription PCR (RT qPCR) also ...

This lead us to question the influence of the innate or cell
autonomouse immune system on early stages of infection. And we thus
settled on an experimental design comparing infections at 5dpi for all

For the parasite oocysts and sporozoites were included as
``environmental'' and ``\textit{in vitro}'' processed lifecycle
stages.

The number of sequence read mappings obtained for individual
replicates ranged from 21,517,055 (sample NMRI\_sporozoites\_rep2) to
593,396,312 (NMRI\_2ndInf\_5dpi\_rep1).

>>>>

  \section*{A dual transcriptomics experiment}

We performed mRNA sequencing of caecum tissue from mice infected with
the apicomplexan parasite \textit{E. falciformis}. Oocysts and
sporozoites were included as ``environmental'' stages and processed
``\textit{in vitro}''. To follow the lifecycle of the parsite, we
compare different timepoints after infection.

We additonally used different mouse strains which displayed different
immunocompetence in infection trials to assess the influence host
immunocompetence on parasite development (figure 1a,b). In these
experiments we analysed RAG-1-/- mice and compared them to the
parental C57BL6 mouse line. RAG knockout mich lack adaptive immune
response du to the absence of mature B and T lymphocytes.  All three
mouse strains were used for infections of both naive and previously
challenge infected animals.

Concerning basic phenotyping, differences in oocyst output were
visible both between the immunocompetent (NMRI and C57BL6) and
immunodeficient (RAG-1-/-) host strain, but also between the two
similarly immunocompetent (NMRI and C57BL6) strains (figure
1a). Surprisingly this finding was consistent in both in primary and
challenge infection. Similarly development of intestinal stages as
assayed by quantitative reverse transcription PCR (RT qPCR) also
... ADD HERE (figure 1b).

We thus aim to analyse the lifecycle of \textit{E. falciformis} under
the influence of (most likely the innate or cell autonomouse) immune
respons at on early stages of infection. We used an experimental
design comparing infections at 5dpi for all possible experimental
conditions (all three mouse strains in naive and challenge infection),
and analyse additional infections at 3dpi (figure 1c).

In our RNAseq experiment we worked with two biological replicates for
all but two conditions (with one and three replicates,
respecitively). For each replicate sample we enriched caecum tissue
from three infected mice for epithelial cells, extracted mRNA and
constructed sequencing libraries. These libraries were sequenced on
several lanes of Illumina GAIIX and HiSeq machines and mapped to both
mouse and parasite genome simulatneously to avoid spurious assignments
of read in ultra conserved genomic regions.

The number of sequence read mappings obtained for individual
replicates ranged from 11,479,260 (sample NMRI\_sporozoites\_rep2) to
145,033,326 (NMRI\_2ndInf\_5dpi\_rep1).
